A studyforrest extension, an annotation of spoken language in the German dubbed movie ``Forrest Gump'' and its audio-description

### 2020-06-02 paper/main.tex
correct typo
introduction: data- vs. model driven, why worth being published, why bother

% data-driven methods
Because naturalistic stimuli are multi-dimensional and their temporal structure
is usually unknown, data-driven methods like intersubject correlation
(ISC)\citep{hasson2004intersubject} or independent component analysis
(ICA)\citep{bartels2004chronoarchitecture} are often used to analyze
corresponding fMRI data. However, data-driven methods alone fall short of
pinpointing their results to specific, time-locked features of a stimulus
\citep{kauttonen2015optimizing}.
% model-driven methods
Model-driven methods, like a general linear model (GLM), can be useful to
investigate how hemodynamic brain responses in isolated regions correlate with
stimulus features. They can explain not just ``how'' the brain is responding to
a stimulus but also ``why'' \citep{hamilton2018revolution}.
% GLM: example studies
For example, studies using GLMs based on annotations of a stimulus' temporal
structure have elucidated how the brain responds to visual features of a movie
\citep{bartels2004mapping} or speech-related features of a narrative
\citep{rocca2019language}.

% We:
For this publication, we revised previous, preliminary annotations done by our
lab and now provide one single annotation with exact onsets/offsets and content
of each sentences, word and phonemes (see Table \ref{tab:overview} for an
overview) spoken in the audio-visual movie ``Forrest Gump''
\citep{ForrestGumpMovie} and its audio-description (i.e. an audio-only variant
of the movie)\citep{ForrestGumpDVD}.
% the studyforrest dataset
fMRI data of participants watching the audio-visual movie
\citep{hanke2016simultaneous} and listening to the audio-description
\citep{hanke2014audiomovie} are the core data of the publicly available
\textit{studyforrest} dataset (\href{www.studyforrest.org}{studyforrest.org}).
% why
This publication enables researchers to model time-locked hemodynamic brain
responses that correlate with different aspects of spoken occurring in a 2-hour
lasting naturalistic stimulus.
% extending previous non-speech annotations
The current annotation of speech extends already available annotations of
portrayed emotions \citep{labs2015portrayed}, perceived emotions
\citep{lettieri2019emotionotopy}, as well as cuts and locations depicted in the
movie \citep{haeusler2016annotation}. All annotations can be used in any study
focusing on aspects of real-life cognition by serving as additional confound
measures describing key properties of major building blocks of the stimulus.

\todo[inline]{Das “why one should model” ist jetzt drin; Frage an dich als
Datenanalyse-Pro: Man kann doch bestimmt auch data-driven methods irgendwie
“constrainen”, und die durch eine Annotation verbessern? Hast du da eine
Formulierung und/oder auch ein (high-level) Paper für mich zu dem Thema?}

@@

\section*{Dataset validation}
% Information about any validation carried out and/or any limitations of the
% datasets, including any allowances made for controlling bias or unwanted
% sources of variability. ~170 words => das wird beim folgenden Schreiben
% definitiv nichts werden intro
In order to assess data quality, we investigated if contrasting speech-related
events to events without speech lead to increased activation in areas known to
be involved in semantic processing \citep{hickok2007cortical}.
% nouns and proper nouns resp. > coordinate junctions
Moreover, we tested if similar linguistic concepts providing high semantic
information (proper nouns and nouns) contrasted to a concept providing low
semantic information (coordinate conjunctions) lead to increased activation in
congruent brain areas.
% data source
% data aligned by applying an affine transformation only are available in
% bold_dico_dico7Tad2grpbold7Tad.nii.gz.
% data aligned by non-linear warping are available in
% bold_dico_dico7Tad2grpbold7Tad_nl.nii.gz.
We used a dataset providing blood oxygenation level-dependent (BOLD) functional
magnetic resonance imaging (fMRI) data of 20 subjects (age 21–38 years, mean age
26.6 years, 12 male) listening to the 2h audio-description (\unit[7]{Tesla},
\unit[2]{s} repetition time, 3599 volumes, 36 axial slices, thickness
\unit[1.4]{mm}, \unit[1.4 $\times$ 1.4]{mm} in-plane resolution, \unit[224]{mm}
field-of-view) \citep{hanke2014audiomovie}.
% motion correction
Data were already corrected for motion at the scanner computer.
% alignment
Further, individual BOLD time-series were already aligned by non-linear warping
to a study-specific T2*-weighted echo planar imaging (EPI) group template (cf.
\citep{hanke2014audiomovie} for details on how the group template was created).
% steps done by C.H.
All further analysis steps were carried out using Feat v6.00 (FMRI Expert
Analysis Tool)\citep{woolrich2001autocorr} as part of FSL v5.0.9 (FMRIB's
Software Library)\citep{smith2004fsl}.
% dropping of subject 10
Data of one participant was dropped to due to invalid distortion correction
during scanning.
% temporal and spatial filter, brain extraction
Data were temporally high-pass filtered (cut-off \unit[150]{s}), spatially
smoothed (Gaussian kernel; \unit[4.0]{mm} FWHM), and the brain was extracted
from surrounding tissue.
% normalization
A grand-mean intensity normalization of the entire 4D dataset was performed by a
single multiplicative factor.

>>>>

\todo[inline]{revised: data-driven vs./in combination (with) model-driven}
% data-driven methods
Because naturalistic stimuli are multi-dimensional and their temporal structure
is usually unknown, data-driven methods like intersubject correlation
(ISC)\citep{hasson2004intersubject} or independent component analysis
(ICA)\citep{bartels2004chronoarchitecture} are often used to analyze
corresponding fMRI data. However, data-driven methods alone fall usually short
of pinpointing their results to specific, time-locked features of a stimulus
\citep{kauttonen2015optimizing}.
% model-driven methods
Model-driven methods, like a general linear model (GLM), can be useful to
investigate how hemodynamic brain responses in isolated regions correlate with
stimulus features. They can explain not just ``how'' the brain is responding to
a stimulus but also ``why'' \citep{hamilton2018revolution}.
% GLM: example studies
For example, studies using GLMs based on annotations of a stimulus' temporal
structure have elucidated how the brain responds to visual features of a movie
\citep{bartels2004mapping} or speech-related features of a narrative
\citep{rocca2019language}.
% combination of data-driven and model-driven approaches
Additionally, stimulus annotations can inform data-driven methods about a
stimulus' temporal dynamics, or model-driven and data-driven methods can be
combined to improve interpretability of results \citep{lahnakoski2012stimulus}.

\todo[inline]{revised: why this anno is better than any previous anno and worth
being published as a paper}
% we:
For this publication, we revised previous, preliminary annotations done by our
lab and now provide one single annotation with exact onsets/offsets and content
of each sentences, word and phonemes (see Table \ref{tab:overview} for an
overview) spoken in the audio-visual movie ``Forrest Gump''
\citep{ForrestGumpMovie} and its audio-description (i.e. an audio-only variant
of the movie)\citep{ForrestGumpDVD}.
% the studyforrest dataset
fMRI data of participants watching the audio-visual movie
\citep{hanke2016simultaneous} and listening to the audio-description
\citep{hanke2014audiomovie} are the core data of the publicly available
\textit{studyforrest} dataset (\href{www.studyforrest.org}{studyforrest.org}).
% why even bother?
This publication enables researchers to model hemodynamic brain responses that
correlate with a variety of time-locked aspects of spoken language ranging from
a speaker's identity, to phonology, grammar, and semantics.
% extending previous non-speech annotations
The current annotation of speech extends already available annotations of
portrayed emotions \citep{labs2015portrayed}, perceived emotions
\citep{lettieri2019emotionotopy}, as well as cuts and locations depicted in the
movie \citep{haeusler2016annotation}. All annotations can be used in any study
focusing on aspects of real-life cognition by serving as additional confound
measures describing key properties of major building blocks of the stimulus.

@@

\section*{Dataset validation}
% Information about any validation carried out and/or any limitations of the
% datasets, including any allowances made for controlling bias or unwanted
% sources of variability. ~170 words => das wird beim folgenden Schreiben
% definitiv nichts werden intro
In order to assess data quality, we investigated if contrasting speech-related
events to events without speech lead to increased activation in areas known to
be involved in semantic processing \citep{hickok2007cortical}.
% nouns and proper nouns resp. > coordinate junctions
Moreover, we tested if similar linguistic concepts providing high semantic
information (proper nouns and nouns) contrasted to a concept providing low
semantic information (coordinate conjunctions) lead to increased activation in
congruent brain areas.
% data source
% data aligned by applying an affine transformation only are available in
% bold_dico_dico7Tad2grpbold7Tad.nii.gz.
% data aligned by non-linear warping are available in
% bold_dico_dico7Tad2grpbold7Tad_nl.nii.gz.
We used a dataset providing blood oxygenation level-dependent (BOLD) functional
magnetic resonance imaging (fMRI) data of 20 subjects (age 21–38 years, mean age
26.6 years, 12 male) listening to the 2h audio-description (\unit[7]{Tesla},
\unit[2]{s} repetition time, 3599 volumes, 36 axial slices, thickness
\unit[1.4]{mm}, \unit[1.4 $\times$ 1.4]{mm} in-plane resolution, \unit[224]{mm}
field-of-view) \citep{hanke2014audiomovie}.
% motion correction
Data were already corrected for motion at the scanner computer.
% alignment
Further, individual BOLD time-series were already aligned by non-linear warping
to a study-specific T2*-weighted echo planar imaging (EPI) group template (cf.
\citep{hanke2014audiomovie} for details on how the group template was created).
% steps done by C.H.
All further analysis steps were carried out using Feat v6.00 (FMRI Expert
Analysis Tool)\citep{woolrich2001autocorr} as part of FSL v5.0.9 (FMRIB's
Software Library)\citep{smith2004fsl}.
% dropping of subject 10
Data of one participant were dropped to due to invalid distortion correction
during scanning.
% temporal and spatial filter, brain extraction
Data were temporally high-pass filtered (cut-off \unit[150]{s}), spatially
smoothed (Gaussian kernel; \unit[4.0]{mm} FWHM), and the brain was extracted
from surrounding tissue.
% normalization
A grand-mean intensity normalization of the entire 4D dataset was performed by a
single multiplicative factor.
